Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal Model to Beyond

### 2024-04-05 main.tex
:pencil: Update abstract

% tabularx, ragged2e
\newcolumntype{L}{>{\RaggedRight}X}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\renewcommand\tabularxcolumn[1]{m{#1}}

@@

% optidef
\DeclareDocumentEnvironment{customopti}{D||{\defaultProblemFormat} O{\defaultConstraintFormat} D<>{} m m m m m}{%
	\ifthenelse{\equal{#3}{b}}{%
		\ifthenelse{\equal{#1}{s}}%
		% Short version problem
		{\setFormatShort{#4}{#5}\BaseMiniStar{#2}{#5}{#6}{#8}{#4}{#3}}%
		% Long version problem
		{\setFormatLong{#4}{#5}\BaseMiniStar{#2}{#5}{#6}{#8}{#4}{#3}}%
	}{%
		\ifthenelse{\equal{#1}{s}}%
		% Short version problem
		{\setFormatShort{#4}{#5}\BaseMini{#2}{#5}{#6}{#7}{#8}{#4}}%
		% Long version problem
		{\setFormatLong{#4}{#5}\BaseMini{#2}{#5}{#6}{#7}{#8}{#4}}%
	}%
}%
{\endBaseMini\toggletrue{bodyCon}}

\DeclareDocumentEnvironment{customopti*}{D||{\defaultProblemFormat} O{\defaultConstraintFormat} D<>{} m m m m m}{%
	\ifthenelse{\equal{#1}{s}}%
	% Short version problem
	{\setFormatShort{#4}{#5}\BaseMiniStar{#2}{#5}{#6}{#8}{#4}{#3}}%
	% Long version problem
	{\setFormatLong{#4}{#5}\BaseMiniStar{#2}{#5}{#6}{#8}{#4}{#3}}%
}{\endBaseMiniStar\toggletrue{bodyCon}}

\DeclareDocumentEnvironment{customopti!}{D||{\defaultProblemFormat} O{\defaultConstraintFormat} D<>{} m m m m m}{%
	\ifthenelse{\equal{#1}{s}}%
	% Short version problem
	{\setFormatShort{#4}{#5}\BaseMiniExclam{#2}{#5}{#6}{#7}{#8}{#4}{#3}}%
	% Long version problem
	{\setFormatLong{#4}{#5}\BaseMiniExclam{#2}{#5}{#6}{#7}{#8}{#4}{#3}}%
}{\endBaseMiniExclam\toggletrue{bodyCon}}

@@

\begin{document}
\title{Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond}
\author{
	\IEEEauthorblockN{
		Yang~Zhao,~\IEEEmembership{Member,~IEEE,}
		Hongyu~Li,~\IEEEmembership{Graduate Student Member,~IEEE,}\\
		Massimo~Franceschetti,~\IEEEmembership{Fellow,~IEEE,}
		and~Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE}
	}
	% \thanks{
	% 	The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{yang.zhao18, b.clerckx\}@imperial.ac.uk).
	% }
}
\maketitle

\begin{abstract}
	This paper investigates how a passive \gls{ris} can reshape the \gls{mimo} point-to-point channel in terms of singular values.
	We depart from the widely-adopted diagonal phase shift model to a general \gls{bd} architecture, which provides superior shaping capability thanks to in-group connections between elements.
	An efficient \gls{rcg} algorithm is tailored for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size, then invoked for the Pareto frontier of channel singular values.
	To understand the gain from off-diagonal entries, we also derive analytical singular value bounds in rank-deficient and fully-connected scenarios.
	As a side product, we tackle \gls{mimo} rate maximization problem by alternating between active beamformer (eigenmode transmission) and passive beamformer (\gls{rcg} algorithm) until convergence.
	A low-complexity suboptimal solution based on channel shaping is also proposed, where the decoupled problem is formulated as channel power maximization and solved in closed form iteratively.
	% We then show how channel shaping decouples the design and amounts to a low-complexity suboptimal solution.
	% A low-complexity suboptimal solution decoupling both blocks is also proposed, where the channel shaping subproblem is formulated as channel power maximization and solved in closed form iteratively.
	Theoretical analysis and numerical evaluation reveal that the shaping advantage of \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions, stemming from stronger subchannel rearrangement and subspace alignment capabilities.
\end{abstract}

\begin{IEEEkeywords}
	Reconfigurable intelligent surface, multi-input multi-output, manifold optimization, singular value control, rate maximization.
\end{IEEEkeywords}

>>>>

@@

@@

\begin{document}
\title{Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal Model to Beyond}
\author{
	\IEEEauthorblockN{
		Yang~Zhao,~\IEEEmembership{Member,~IEEE,}
		Hongyu~Li,~\IEEEmembership{Graduate Student Member,~IEEE,}\\
		Massimo~Franceschetti,~\IEEEmembership{Fellow,~IEEE,}
		and~Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE}
	}
	% \thanks{
	% 	The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{yang.zhao18, b.clerckx\}@imperial.ac.uk).
	% }
}
\maketitle

\begin{abstract}
	This paper investigates how a passive \gls{ris} can manipulate the \gls{mimo} point-to-point channel in terms of singular values and power gain.
	We depart from the widely-adopted diagonal phase shift model to a \gls{bd} architecture, which enables \emph{cooperative} wave scattering thanks to in-group connections between elements.
	An efficient and universal asymmetric \gls{bd}-\gls{ris} design framework is proposed over group-wise \emph{geodesic} \gls{rcg} update in the Stiefel manifold.
	For the singular value shaping problem, we characterize their Pareto frontiers by weighted sum optimization via geodesic \gls{rcg}, which generalizes most relevant metrics (e.g., degree of freedom and condition number).
	Many useful analytical singular value bounds are also derived in rank-deficient and fully-connected (a.k.a. unitary) scenarios.
	% For the singular value problem, we consider a numerical optimization approach that characterizes the Pareto frontiers as well as an analytical bounding approach in special (i.e., rank-deficient and fully-connected) scenarios.
	For the power gain maximization problem, we provide a closed-form iterative solution by group-wise \gls{svd} that is even faster than geodesic \gls{rcg}.
	% Those channel shaping designs have a two-fold advantage.
	Those channel shaping designs have two major implications.
	First, they demystify the performance gain from off-diagonal entries in terms of \emph{subchannel rearrangement} and \emph{subspace alignment}.
	% help to demystify the gain from off-diagonal entries and reveal the shaping advantage of \gls{bd}-\gls{ris} in subchannel rearrangement and subspace alignment.
	Second, they inspire a two-stage suboptimal solution to decouple the joint active and passive beamforming design.
	% \gls{ris}-transceiver optimization and suggest a low-complexity suboptimal solution based on channel power maximization.
	As a side product, we tackle the \gls{bd}-\gls{ris}-aided \gls{mimo} rate maximization problem by a local-optimal \gls{ao} approach and a two-stage shaping-inspired approach.
	Simulation results indicate that the rate deficit from the latter diminishes as the \gls{bd}-\gls{ris} evolves towards unitary, while its shaping advantages over diagonal \gls{ris} increase \gls{mimo} dimensions and decrease with spatial correlation.

	% , and .

	% (i) , and (ii) the gap between the two approaches vanishes as \gls{bd}-\gls{ris} evolves from diagonal to unitary.
	% not only demystify the advantage of \gls{bd}-\gls{ris} in subchannel rearrangement and subspace alignment, but also decouple the joint \gls{ris}-transceiver optimization.
	% Those channel shaping designs not only demystify the advantage of \gls{bd}-\gls{ris} in subchannel rearrangement and subspace alignment, but also decouple the joint \gls{ris}-transceiver optimization.

	% geodesic \gls{rcg} algorithm is tailored for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size, then invoked for the Pareto frontier of channel singular values.
	% To understand the gain from off-diagonal entries, we also derive analytical singular value bounds in rank-deficient and fully-connected scenarios.
	% As a side product, we tackle \gls{mimo} rate maximization problem by alternating between active beamformer (eigenmode transmission) and passive beamformer (\gls{rcg} algorithm) until convergence.
	% A low-complexity suboptimal solution based on channel shaping is also proposed, where the decoupled problem is formulated as channel power maximization and solved in closed form iteratively.
	% % We then show how channel shaping decouples the design and amounts to a low-complexity suboptimal solution.
	% % A low-complexity suboptimal solution decoupling both blocks is also proposed, where the channel shaping subproblem is formulated as channel power maximization and solved in closed form iteratively.
	% Theoretical analysis and numerical evaluation reveal that the shaping advantage of \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions, stemming from stronger subchannel rearrangement and subspace alignment capabilities.
\end{abstract}

\begin{IEEEkeywords}
	Reconfigurable intelligent surface, multi-input multi-output, manifold optimization, channel shaping, rate maximization.
\end{IEEEkeywords}
